---
title: "GitHub Repository Fork Network"
author: "Jesse Abdul"
date: "12/21/2024"
output:
  html_document:
    code_folding: hide
    df_print: paged
  html_notebook: 
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_chunk$set(echo = TRUE) # Display code in the output document
opts_chunk$set(comment="")  # Get rid of ## on output knit to html. 
library(tibble) # Import first so it does not mask igraph simplify
library(igraph)
library(jpeg) 
library(png)
library(moments) # install this package to compute 2nd moment easily 
source("Utility/reify_link_communities.R")
source("Utility/degree_domain.R")
source("Utility/nonzero_degree_distribution.R")
source("Utility/top_n_vertices.R")
```

____________________________________________________________
**Read in the generated network graph and calculate metrics to annotate the graph**
```{r}
GFN <- read_graph("Networks/github_forked_repo_network.graphml", format="graphml")
GFN$name <- "GitHub Repository Fork Network"

V(GFN)$in_degree <- degree(GFN, mode="in")
V(GFN)$out_degree <- degree(GFN, mode="out")
V(GFN)$page_rank <- page_rank(GFN)$vector
V(GFN)$betweenness <- betweenness(GFN, normalized=TRUE)
V(GFN)$harmonic_centrality  <- harmonic_centrality(GFN, normalized=TRUE)


# calculate graph-level metrics
GFN$global_transitivity <- transitivity(GFN, type="global")
GFN$diameter <- diameter(GFN)

# calculate the louvain communities using the modularity metric (use a temporarily generated undirected graph)
V(GFN)$comm_louvain <- membership(cluster_louvain(as_undirected(GFN, mode="collapse")))

# calculate the infomap communities using the modularity metric
V(GFN)$comm_infomap <- membership(cluster_infomap(GFN))

# disconnected graph, so assortativity is NaN
# GFN$deg_assortativity <- assortativity_degree(GFN)
GFN$owner_type_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$owner_type))
GFN$in_degree_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$in_degree))
GFN$out_degree_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$out_degree))
GFN$betweenness_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$betweenness))
GFN$page_rank_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$page_rank))
GFN$harmonic_centrality_assortativity <- assortativity_nominal(GFN, types=as.factor(V(GFN)$harmonic_centrality))


GFN$mean_in_degree = mean(V(GFN)$in_degree)
GFN$mean_out_degree = mean(V(GFN)$out_degree)
E(GFN)$weight = 1;

summary(GFN)

```



**Show the graph-level metrics**
```{r}
tibble(
  global_transitivity = GFN$global_transitivity,
  diameter = GFN$diameter,
  mean_in_degree = GFN$mean_in_degree,
  mean_out_degree = GFN$mean_out_degree,
  owner_type_assortativity = GFN$owner_type_assortativity,
  in_degree_assortativity = GFN$in_degree_assortativity,
  out_degree_assortativity = GFN$out_degree_assortativity,
  betweenness_assortativity = GFN$betweenness_assortativity,
  page_rank_assortativity = GFN$page_rank_assortativity,
  harmonic_centrality_assortativity = GFN$harmonic_centrality_assortativity
)
```

**calculate the standard metrics used to predict regimes in random networks**
```{r}
GFN_degree <- degree(GFN) # compute once for efficiency and convenience 
tibble(
 "N = |V|" = vcount(GFN),
 "M = |E|" = ecount(GFN),
 "<k>"     = mean(GFN_degree),
 "ln N"    = log(vcount(GFN)),
 "ln N / ln <k>" = log(vcount(GFN))/log(mean(GFN_degree)),
 "ln ln N" = log(log(vcount(GFN)))
)
```


**Generate the random graphs**
```{r}
# completely random graph
GFN_gnm <- sample_gnm(vcount(GFN), ecount(GFN), directed=TRUE)
GFN_gnm$name <- paste("GFN", GFN_gnm$name)
summary(GFN_gnm)

# configuration model
GFN_config <- sample_degseq(V(GFN)$out_degree, V(GFN)$in_degree, method="configuration")
GFN_config$name <- paste("GFN", GFN_config$name)
summary(GFN_config)

# degree preserving randomization
GFN_rewire <- rewire(GFN, keeping_degseq(loops=FALSE, niter=100*ecount(GFN)))
GFN_rewire$name <- "GFN Rewired"
summary(GFN_rewire)

```

**compare the natural network to comparable random networks**
```{r}
GFN_models <- list(GFN, GFN_gnm, GFN_config, GFN_rewire) # must list not c
tibble(
Model         = c("GitHub Fork Network", "G(n,m)", "Config", "Rewire"), 
 strong_components    = sapply(GFN_models, count_components, mode="strong"), # Number of strong components
 weak_components    = sapply(GFN_models, count_components, mode="weak"), # Number of weak components
 percent_giant = sapply(sapply(GFN_models, largest_component), vcount)/sapply(GFN_models, vcount), # Percent in giant component 
 meandist      = sapply(GFN_models, mean_distance, directed=TRUE), # Mean distance 
 transitivity  = sapply(GFN_models, transitivity, type="global"), # Global transitivity  
 degreeassort  = sapply(GFN_models, assortativity_degree, directed=TRUE), # Degree assortativity 
reciprocity    = sapply(GFN_models, reciprocity, mode="default")
)
```



**Show the top N vertices in each vertex metric **
```{r}

tibble(
  In_Degree = top_n_vertices(GFN, V(GFN)$in_degree, 30)$repo_full_name,
  In_Degree_Value = top_n_vertices(GFN, V(GFN)$in_degree, 30)$in_degree,
  Fork_Depth = top_n_vertices(GFN, V(GFN)$fork_depth, 30)$repo_full_name,
  Fork_Depth_Parent = top_n_vertices(GFN, V(GFN)$fork_depth, 30)$fork_depth_parent,
  Fork_Depth_Value = top_n_vertices(GFN, V(GFN)$fork_depth, 30)$fork_depth,
  Betweenness = top_n_vertices(GFN, V(GFN)$betweenness, 30)$repo_full_name,
  Betweenness_Value = top_n_vertices(GFN, V(GFN)$betweenness, 30)$betweenness,
  Page_Rank = top_n_vertices(GFN, V(GFN)$page_rank, 30)$repo_full_name,
  Page_Rank_Value = top_n_vertices(GFN, V(GFN)$page_rank, 30)$page_rank,
  Harmonic_Centrality = top_n_vertices(GFN, V(GFN)$harmonic_centrality, 30)$repo_full_name,
  Harmonic_Centrality_Value = top_n_vertices(GFN, V(GFN)$harmonic_centrality, 30)$harmonic_centrality
  
  
)


```


 
**Plot the in-degree distribution**
```{r}
# display a histogram of the degree distribution
plot(degree_domain(GFN, "in"), nonzero_degree_distribution(GFN, "in"), main="GitHub Fork Network In-Degree Distribution", xlab="k", ylab="p(k)", log="xy"  )

legend("topright", 
       legend= c(paste("k-min =", min(V(GFN)$in_degree)), paste("<k> =", round(moment(V(GFN)$in_degree, order=1), 4)), paste("<k^2> =", round(moment(V(GFN)$in_degree, order=2), 4)), paste("k-max =", max(V(GFN)$in_degree))), bty="n", cex=1.0)


```
**Write the annotated github fork network graph**
```{r}
write_graph(GFN, "Networks/github_forked_repo_network_annotated.graphml", format="graphml")

```
